The Title

### 2020-12-14 resources/paper_calenviro.Rmd
modification to writeup
first models

Indicators classified as pollution 'exposures' in CES includes ozone, measured as the amount of daily maximum 8-hour ozone concentration between May-October, averaged over 2012-14; PM2.5 (fine particulate matter, ~2.5 microns), measured as an annual mean concentration between 2012-14; particulate matter from on- and off-road diesel exhaust, measured as a spatial distribution of diesel PM emissions on a July day in 2012; an index of drinking water contaminants (including nitrate, perchlorate, and animal waste), measured between 2005-13; pesticides, measured as total pounds of selected active pesticide ingredients used in production-agriculture per square mile in the census tract, averaged over 2012-14; toxic releases, measured as toxicity-weighted concentrations of modeled chemical releases to air from facility emissions and off-site incineration, averaged over 2011-13; traffic, measured as vehicle-kilometers per hour per road length, within 150 meters of the census tract boundary in 2013.

@@

Barker, D. J. P., J. G. Eriksson, T. Forsén, and C. Osmond. 2002. “Fetal Origins of Adult Disease: Strength of Effects and Biological Basis.” International Journal of Epidemiology 31(6):1235–39. doi: 10.1093/ije/31.6.1235.
Bhutta, Adnan T., Mario A. Cleves, Patrick H. Casey, Mary M. Cradock, and K. J. S. Anand. 2002. “Cognitive and Behavioral Outcomes of School-Aged Children Who Were Born Preterm: A Meta-Analysis.” JAMA 288(6):728–37. doi: 10.1001/jama.288.6.728.
Ghosh, Jo Kay C., Michelle Wilhelm, Jason Su, Daniel Goldberg, Myles Cockburn, Michael Jerrett, and Beate Ritz. 2012. “Assessing the Influence of Traffic-Related Air Pollution on Risk of Term Low Birth Weight on the Basis of Land-Use-Based Regression Models and Measures of Air Toxics.” American Journal of Epidemiology 175(12):1262–74. doi: 10.1093/aje/kwr469.
Greenfield, Ben K., Jayant Rajan, and Thomas E. McKone. 2017. “A Multivariate Analysis of CalEnviroScreen: Comparing Environmental and Socioeconomic Stressors versus Chronic Disease.” Environmental Health 16. doi: 10.1186/s12940-017-0344-z.
Harley, Kim G., Karen Huen, Raul Aguilar Schall, Nina T. Holland, Asa Bradman, Dana Boyd Barr, and Brenda Eskenazi. 2011. “Association of Organophosphate Pesticide Exposure and Paraoxonase with Birth Outcome in Mexican-American Women.” PLOS ONE 6(8):e23923. doi: 10.1371/journal.pone.0023923.
Laurent, Olivier, Jun Wu, Lianfa Li, Judith Chung, and Scott Bartell. 2013. “Investigating the Association between Birth Weight and Complementary Air Pollution Metrics: A Cohort Study.” Environmental Health 12(1):18. doi: 10.1186/1476-069X-12-18.
Lu, Michael C., and Neal Halfon. 2003. “Racial and Ethnic Disparities in Birth Outcomes: A Life-Course Perspective.” Maternal and Child Health Journal 7(1):13–30. doi: 10.1023/A:1022537516969.
Martinson, Melissa L., and Nancy E. Reichman. 2016. “Socioeconomic Inequalities in Low Birth Weight in the United States, the United Kingdom, Canada, and Australia.” American Journal of Public Health 106(4):748–54. doi: 10.2105/AJPH.2015.303007.
Nepomnyaschy, Lenna, and Nancy E. Reichman. 2006. “Low Birthweight and Asthma Among Young Urban Children.” American Journal of Public Health 96(9):1604–10. doi: 10.2105/AJPH.2005.079400.
Rodriquez, Matthew, and Lauren Zeise. 2017. CalEnviroScreen 3.0 Sourcebook. Sacramento, CA: California Office of Environmental Health Hazard Assessment.
Sims, Mario, Tammy L. Sims, and Marino A. Bruce. 2008. “Race, Ethnicity, Concentrated Poverty, and Low Birth Weight Disparities.” Journal of National Black Nurses’ Association : JNBNA 19(1):12–18.
World Health Organization (WHO). 2018. “Preterm Birth.” Retrieved October 19, 2020 (https://www.who.int/news-room/fact-sheets/detail/preterm-birth).

>>>>

Indicators classified as pollution 'exposures' in CES includes ozone, measured as the amount of daily maximum 8-hour ozone concentration between May-October (parts per million), averaged over 2012-14; PM2.5 (fine particulate matter, ~2.5 microns), measured as an annual mean concentration between 2012-14; particulate matter from on- and off-road diesel exhaust, measured as a spatial distribution of diesel PM emissions on a July day in 2012; an index of drinking water contaminants (including nitrate, perchlorate, and animal waste), measured between 2005-13; pesticides, measured as total pounds of selected active pesticide ingredients used in production-agriculture per square mile in the census tract, averaged over 2012-14; toxic releases, measured as toxicity-weighted concentrations of modeled chemical releases to air from facility emissions and off-site incineration, averaged over 2011-13; traffic, measured as vehicle-kilometers per hour per road length, within 150 meters of the census tract boundary in 2013.

@@

Barker, D. J. P., J. G. Eriksson, T. Forsén, and C. Osmond. 2002. “Fetal Origins of Adult Disease: Strength of Effects and Biological Basis.” International Journal of Epidemiology 31(6):1235–39. doi: 10.1093/ije/31.6.1235.
Bhutta, Adnan T., Mario A. Cleves, Patrick H. Casey, Mary M. Cradock, and K. J. S. Anand. 2002. “Cognitive and Behavioral Outcomes of School-Aged Children Who Were Born Preterm: A Meta-Analysis.” JAMA 288(6):728–37. doi: 10.1001/jama.288.6.728.
Ghosh, Jo Kay C., Michelle Wilhelm, Jason Su, Daniel Goldberg, Myles Cockburn, Michael Jerrett, and Beate Ritz. 2012. “Assessing the Influence of Traffic-Related Air Pollution on Risk of Term Low Birth Weight on the Basis of Land-Use-Based Regression Models and Measures of Air Toxics.” American Journal of Epidemiology 175(12):1262–74. doi: 10.1093/aje/kwr469.
Greenfield, Ben K., Jayant Rajan, and Thomas E. McKone. 2017. “A Multivariate Analysis of CalEnviroScreen: Comparing Environmental and Socioeconomic Stressors versus Chronic Disease.” Environmental Health 16. doi: 10.1186/s12940-017-0344-z.
Harley, Kim G., Karen Huen, Raul Aguilar Schall, Nina T. Holland, Asa Bradman, Dana Boyd Barr, and Brenda Eskenazi. 2011. “Association of Organophosphate Pesticide Exposure and Paraoxonase with Birth Outcome in Mexican-American Women.” PLOS ONE 6(8):e23923. doi: 10.1371/journal.pone.0023923.
Laurent, Olivier, Jun Wu, Lianfa Li, Judith Chung, and Scott Bartell. 2013. “Investigating the Association between Birth Weight and Complementary Air Pollution Metrics: A Cohort Study.” Environmental Health 12(1):18. doi: 10.1186/1476-069X-12-18.
Lu, Michael C., and Neal Halfon. 2003. “Racial and Ethnic Disparities in Birth Outcomes: A Life-Course Perspective.” Maternal and Child Health Journal 7(1):13–30. doi: 10.1023/A:1022537516969.
Martinson, Melissa L., and Nancy E. Reichman. 2016. “Socioeconomic Inequalities in Low Birth Weight in the United States, the United Kingdom, Canada, and Australia.” American Journal of Public Health 106(4):748–54. doi: 10.2105/AJPH.2015.303007.
Nepomnyaschy, Lenna, and Nancy E. Reichman. 2006. “Low Birthweight and Asthma Among Young Urban Children.” American Journal of Public Health 96(9):1604–10. doi: 10.2105/AJPH.2005.079400.
Ponce, Ninez A., Katherine J. Hoggatt, Michelle Wilhelm, and Beate Ritz. 2005. “Preterm Birth: The Interaction of Traffic-Related Air Pollution with Economic Hardship in Los Angeles Neighborhoods.” American Journal of Epidemiology 162(2):140–48. doi: 10.1093/aje/kwi173.
Rodriquez, Matthew, and Lauren Zeise. 2017. CalEnviroScreen 3.0 Sourcebook. Sacramento, CA: California Office of Environmental Health Hazard Assessment.
Sims, Mario, Tammy L. Sims, and Marino A. Bruce. 2008. “Race, Ethnicity, Concentrated Poverty, and Low Birth Weight Disparities.” Journal of National Black Nurses’ Association : JNBNA 19(1):12–18.
World Health Organization (WHO). 2018. “Preterm Birth.” Retrieved October 19, 2020 (https://www.who.int/news-room/fact-sheets/detail/preterm-birth).
